---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dev = "ragg_png",
  dpi = 600,
  fig.width = 12,
  fig.height = 8,
  message = FALSE,
  warning = FALSE
)

```


# ScanCentricPeakCharacterization


The goal of the ScanCentricPeakCharacterization (SCPC) package is to facilitate scan-centric, frequency based, peak characterization of profile level, multi-scan, direct-injection Fourier-transform mass spectrometry data.

You can read more about the merits of this scan-centric method in:

RM Flight, JM Mitchell & HNB Moseley, "Scan-Centric, Frequency-Based Method for Characterizing Peaks from Direct Injection Fourier transform Mass Spectrometry Experiments", Metabolites 2022, 12(6), 515; https://doi.org/10.3390/metabo12060515

## License

This package is licensed with a [BSD-like license](LICENSE.md) with a 4th clause: **No commercial use.**

Academics who want to use it at their institution, please try it.

If you are at a business / for-profit and want to use it, please contact the authors (Robert Flight, rflight79 at gmail; Hunter Moseley, hunter dot moseley at gmail) about licensing.
Please contact us even if you aren't sure what would be required for licensing, we do want people to use it.

## Installation

You can install SCPC from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("MoseleyBioinformaticsLab/ScanCentricPeakCharacterization")
```

## Documentation Site

You can browse the documentation online [here](https://moseleybioinformaticslab.github.io/ScanCentricPeakCharacterization).

## Setup

```{r setup}
library(ScanCentricPeakCharacterization)
library(dplyr)
library(patchwork)
library(ggplot2)
#theme_set(cowplot::theme_cowplot())
```

## Theory

### Converting m/z to Frequency

Outside of the scan-centric nature of this peak-characterization, the second most important feature is the conversion from m/z to frequency.
This is done to make evenly spaced data.
If you acquire Orbitrap / ICR type mass spectrometry data over any decent range, there is an increasing spacing between individual m/z points.

We will load up an example direct-injection lipidomics sample acquired on a Thermo-Fisher Fusion instrument to demonstrate.

```{r example_lipidomics}
mzml_lipid = SCMzml$new(system.file("extdata/lipid_example.mzML", package = "ScanCentricPeakCharacterization"))
mzml_lipid$extract_mzml_data()
mzml_lipid$predict_frequency()
```


```{r show_spacing}
mzml_lipid$mzml_df_data[[1]] %>%
  dplyr::filter(convertable) %>%
  ggplot(aes(x = mz, y = mean_offset)) +
  geom_point()
```

We can see here that the difference or offset of m/z points is increasing with m/z.

In contrast, frequency is defined as the difference over m/z, and therefore is constant.

$$mz_{mean} = mean(mz_{p1}, mz_{p2})$$

$$mz_{diff} = mz_{p2} - mz_{p1}$$

$$frequency = \frac{mz_{mean}}{mz_{diff}}$$

```{r show_frequency}
mzml_lipid$mzml_df_data[[1]] %>%
  dplyr::filter(convertable) %>%
  ggplot(aes(x = mean_frequency, y = mean_freq_diff)) +
  geom_point()
```

However, we can more generally define the conversion of m/z to frequency using a linear model of the form:

$$frequency = a + \frac{x}{mz} +  \frac{y}{\sqrt{mz}} + \frac{z}{\sqrt[3]{mz}}$$

And we can verify that with a plot of the m/z vs frequency and their predicted values, in a couple of ways, as well as a plot of the residuals.

```{r show_predictions}
mzml_lipid$check_frequency_model()
```

See the example of `SCMzml` below to see how we can change the model being used.

## Basic Objects and Classes

A note about the assumptions baked into SCPC:

* Data is acquired in multiple, direct-injection scans (i.e. there is no chromatography component).
* The scans are not aggregated together.
* The data is in **profile mode**, not centroided.
* The M/Z range for each scan is identical.
* The data is in an open format supported by `MSnbase` and `mzR`.

### SCMzml

We assume `mzML` is the most likely mass-spectrometer format, and `SCMzml` is an R6 container around the `mzML` file (although it should work with **any** of the formats supported by `MSnbase`).
It has basic functionality for loading the `mzML` data using `MSnbase`, filtering out scans that are not required, and converting the M/Z to frequency space for subsequent peak detection.

It is good to check a few samples using `SCMzml` before peak characterization.
This is especially important if you need to define custom scan filtering and choosing a single frequency model before attempting full peak characterization.

Let's go through some of it using the example file included with SCPC.
This sample is a lipidomics sample acquired on a Thermo-Fisher tribrid Orbitrap Fusion over 15 minutes, with a combination of MS1 and MS2 scans, acquired in profile mode.
It has been filtered to a subset of M/Z and total scans to save space.
The first 7.5 minutes are the MS1 primary scans, and the remaining 7.5 minutes are combination of MS1 and MS2 scans of the highest intensity M/Z in the MS1 primary scans.

#### Load the Data

```{r}
#| label: load-data
lipid_sample = system.file("extdata", "lipid_example.mzML", package = "ScanCentricPeakCharacterization")
sc_mzml = SCMzml$new(lipid_sample)
```

After instantiating the object, we extract the data into a list of data.frames that make it easier to work with.
This may change over time to use more efficient data structures.

```{r}
#| label: extract-mzml
sc_mzml$extract_mzml_data()
```

#### Examine the Scan Level Information

We can see what is reported at the scan level:

```{r}
#| label: show-scans
head(sc_mzml$scan_info) |>
  gt::gt()
```

And in the data.frame created from the mzML data (here is a single scan):

```{r}
#| label: show-data
head(sc_mzml$mzml_df_data[[1]]) |>
  gt::gt()
```

#### Frequency and M/Z Models

As discussed above (see [Theory](#theory)), SCPC is based on the idea that we should work in a pseudo-frequency space, which is achieved by calculating an initial frequency based on the M/Z spacing, and then fit that to a model based on the geometry of the Orbitrap or ICR.

For Thermo Orbitrap (Fusion and Fusion Lumos), that our lab primarily uses, this model looks like:

$$frequency = a + \frac{x}{mz} +  \frac{y}{\sqrt{mz}} + \frac{z}{\sqrt[3]{mz}}$$

In contrast, for Bruker SolariX, the model has one fewer term:

$$frequency = a + \frac{x}{mz} +  \frac{y}{\sqrt{mz}}$$

Note how these models are encoded, the (-) indicates a fraction, and the actual fraction indicates a root (either square or cube in these cases):

```{r}
#| label: example-models
# thermo fusion model:
thermo = c("a.freq" = 0, "x.freq" = -1, "y.freq" = -1/2, "z.freq" = -1/3)

# bruker
bruker = c("a.freq" = 0, "x.freq" = -1, "y.freq" = -1/2)
```

```{r}
#| label: predict-frequency
sc_mzml$predict_frequency()
```

Here is the scan information after prediction, it now has model terms related to the above equations:

```{r}
sc_mzml$scan_info |>
  head() |>
  gt::gt()
```

And the mz data.frame now has frequency related data as well:

```{r}
sc_mzml$mzml_df_data[[1]] |>
  head() |>
  gt::gt()
```

We can see a bunch more information added to the scan level summary data and the M/Z data.frame after we predict frequency in each scan.

We can check the frequency fit using `check_frequency`.

```{r}
#| label: check-frequency
sc_mzml$check_frequency_model()
```

What we've plotted are for a single scan:

* The M/Z and the calculated frequency (black), and the fit line according to the model above;
* The calculated frequency from M/Z points, and the predicted frequency after fitting a model, as well as the 1 - 1 line (red);
* Calculated - predicted frequency as a function of M/Z

For all scans, we have the median and maximum absolute deviation (MAD) of the calculated - predicted residuals.

#### Changing the Frequency Model

We can change the frequency model and recheck it easily.
Let's use a model where we change the square root to a cube root.

```{r}
#| label: change-model
correct_model = c("a.freq" = 0, "x.freq" = -1, "y.freq" = -1/2, "z.freq" = -1/3)
off_model = c("a.freq" = 0, "x.freq" = -1, "y.freq" = -1/3)

sc_mzml$frequency_fit_description = off_model
sc_mzml$predict_frequency()
sc_mzml$check_frequency_model()
```

From this diagram, we are pretty sure that the model is incorrect for our instrument.

Before proceeding, let's revert it.

```{r}
#| label: revert-model
sc_mzml$frequency_fit_description = correct_model
sc_mzml$predict_frequency()
```


#### Filter Scans

Next we set up our filter scans function.
Let's plot the scans first.

```{r}
#| label: scan-info-plot
sc_mzml$scan_info |>
  ggplot(aes(x = rtime, xend = rtime,
             y = 0, yend = tic,
             color = y.freq)) +
  geom_segment()
```

We've plotted the scans by their *retention time* (rtime), and their height is the *total intensity chromatogram* (tic), and colored by the value of the *y frequency term* (y.freq).

For **this** data, we only want those scans below an rtime of 450, and with a y.freq >= 2.9e7.
After using these criteria, the default function also uses `boxplot.stats` on the y.freq term to check for possible outlier scans.

```{r}
#| label: create-filter-scans
sc_mzml$generate_filter_scan_function(rtime = c(NA, 450),
                                      y.freq = c(2.9e7, NA))
sc_mzml$filter_scans()
```

Now we can see which scans are being excluded with this filter.

```{r}
#| label: filter-scans-table
sc_mzml$scan_info |>
  head() |>
  gt::gt()
```

```{r}
#| label: filter-scans-plot
sc_mzml$scan_info |>
  ggplot(aes(x = rtime, xend = rtime,
             y = 0, yend = tic,
             color = keep)) +
  geom_segment()
```

Changing the filters here using variations of `rtime` and `y.freq` should be simple.
If you have more involved needs, you can write your own filtering function.
`filter_scans_builtin` is another example of a function you can use as a template.

If you want to use a custom function (named *my_custom_filter*), you can add it like this:

```{r}
#| label: filter-scans-custom
#| eval: false
sc_mzml$generate_filter_scan_function(f_function = my_custom_filter)
```

#### Choose Frequency Model

After filtering scans, we need to pick a single frequency model.
Our [publication](https://doi.org/10.3390/metabo12060515) showed using each scan's own model is a bad idea, and we don't advise using some conglomeration of models either.
Instead, the default is to take the model with the y-term closest to the median of all the y-terms.

```{r}
#| label: choose-model
sc_mzml$generate_choose_frequency_model_function()
sc_mzml$choose_frequency_model()
sc_mzml$frequency_coefficients
```

#### Convert to Frequency

Finally, we can convert our M/Z data to frequency for use in peak characterization.

```{r}
#| label: convert-frequency
sc_mzml$convert_to_frequency()
```

### SCCharacterizePeaks

`SCCharacterizePeaks` controls the overall interplay between:

* the `SCZip` container that will hold the original and final data;
* the `SCMzml` object that loads mzml data, transforms it to frequency space, and filters out scans that don't seem to belong;
* the `SCPeakRegion` and `SCPeakRegionFinder` that actually do all of the peak characterization.

Let's give an example using an example lipid file.
This chunk is **not** evaluated due to peak characterization taking so long.

```{r example_char}
lipid_sample = system.file("extdata", "lipid_example.mzML", package = "ScanCentricPeakCharacterization")
sc_char = SCCharacterizePeaks$new(lipid_sample, out_file = here::here("lipid_sample.zip"))
sc_char$load_file()
```

```{r}
#| label: run-all
#| eval: false
sc_char$generate_filter_scan_function(rtime = c(NA, 450),
                                      y.freq = c(2.9e7, NA))
sc_char$generate_choose_frequency_model_function()
sc_char$prepare_mzml_data()
sc_char$find_peaks()
sc_char$summarize()
sc_char$save_peaks()
sc_char$write_zip()
```

#### Run Everything

If you have a large number of samples to run, and you know you will be using the same functions for filtering scans and choosing the frequency model, you can use the `run_all` instead.
It takes two arguments, the `filter_scan_function`, and the `choose_frequency_model_function`.

```{r run_everything, eval = FALSE}
# not run
sc_char = SCCharacterizePeaks$new("file.mzML", out_file = "file.zip")
sc_char$run_all(filter_scan_function = custom_filter,
                choose_frequency_model_function = custom_frequency_model)
```

You should see the vignette on custom functions if you want to go this route.

### SCPeakRegionFinder

`SCPeakRegionFinder` is similar to `SCCharacterizePeaks` in that it is more of a controlling workflow object.
It serves to coordinate all the steps that need to happen for peak characterization outside of the conversion to frequency, which is the purview of the `SCMzml` object.
The `SCPeakRegionFinder` acts on the `SCPeakRegions` object, which has all of the data and methods.

### SCPeakRegions

`SCPeakRegions` holds the frequency data and the methods.
It is controlled by `SCPeakRegionFinder`.

### SCZip

We wanted a fairly generic way to store the original mzML file, any metadata generated about it, the binary output of `SCPeakRegionFinder`, and a JSONized peak list that can be used for assignment.
What we decided on was a simple zip file that keeps those objects together.
When we create a new `SCZip`, we actually create a temp directory, and move all the data there, and unzip it so that it is easily accessible and pieces can be modified easily.

For example, we can see where the temp data lives for our previously created `sc_char` object.

```{r}
#| label: show-temp
sc_char$temp_loc
dir(sc_char$temp_loc)
```
