---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE
)

```


# ScanCentricPeakCharacterization


The goal of the ScanCentricPeakCharacterization package is to facilitate scan-centric peak characterization of profile level, multi-scan, direct-injection Fourier-transform mass spectrometry data.

You can read more about the merits of this scan-centric method in:

RM Flight, JM Mitchell & HNB Moseley, "Scan-Centric, Frequency-Based Method for Characterizing Peaks from Direct Injection Fourier transform Mass Spectrometry Experiments", bioRxiv, doi: https://doi.org/10.1101/2022.04.14.488423

## License

This package is licensed with a [BSD-like license](LICENSE.md) with a 4th clause: No commercial use.

Academics who want to use it at their institution, please try it.

If you are at a business / for-profit and want to use it, please contact the authors about licensing.
Please contact us even if you aren't sure what would be required for licensing, we do want people to use it.

## Installation

You can install ScanCentricPeakCharacterization from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("MoseleyBioinformaticsLab/ScanCentricPeakCharacterization")
```

## Setup

```{r setup}
library(ScanCentricPeakCharacterization)
library(dplyr)
library(patchwork)
library(ggplot2)
theme_set(cowplot::theme_cowplot())
```

## Theory

### Converting m/z to Frequency

Outside of the scan-centric nature of this peak-characterization, the second most important feature is the conversion from m/z to frequency.
This is done to make evenly spaced data.
If you acquire Orbitrap / ICR type mass spectrometry data over any decent range, their is an increasing spacing between individual m/z points.

We will load up an example direct-injection lipidomics sample acquired on a Thermo-Fisher Fusion instrument to demonstrate.

```{r example_lipidomics}
raw_lipid = SCRaw$new(system.file("extdata/lipid_example.mzML", package = "ScanCentricPeakCharacterization"))
raw_lipid$extract_raw_data()
raw_lipid$predict_frequency()
```


```{r show_spacing}
raw_lipid$raw_df_data[[1]] %>%
  dplyr::filter(convertable) %>%
  ggplot(aes(x = mz, y = mean_offset)) +
  geom_point()
```

We can see here that the difference or offset of m/z points is increasing with m/z.

In contrast, frequency is defined as the difference over m/z, and therefore is constant.

$$mz_{mean} = mean(mz_{p1}, mz_{p2})$$

$$mz_{diff} = mz_{p2} - mz_{p1}$$

$$frequency = \frac{mz_{diff}}{mz_{mean}}$$

```{r show_frequency}
raw_lipid$raw_df_data[[1]] %>%
  dplyr::filter(convertable) %>%
  ggplot(aes(x = mean_frequency, y = mean_freq_diff)) +
  geom_point()
```

However, we can more generally define the conversion of m/z to frequency using a linear model of the form:

$$frequency = a + \frac{y}{\sqrt{mz}} + \frac{z}{\sqrt[3]{mz}}$$

And we can verify that with a plot of the m/z vs frequency and their predicted values, in a couple of ways, as well as a plot of the residuals.

```{r show_predictions}
raw_lipid$check_frequency_model()
```

See the example of `SCRaw` below to see how we can change the model being used.

## Basic Objects and Classes

### SCCharacterizePeaks

`SCCharacterizePeaks` controls the overall interplay between:

* the `SCZip` container that will hold the original and final data;
* the `SCRaw` object that loads raw data, transforms it to frequency space, and filters out scans that don't seem to belong;
* the `SCPeakRegion` and `SCPeakRegionFinder` that actually do all of the peak characterization.

It also has the ability to do the first preparation steps of the data and check that the frequency model is appropriate.

Let's give an example using an example lipid file.

```{r example_char}
lipid_sample = system.file("extdata", "lipid_example.mzML", package = "ScanCentricPeakCharacterization")
sc_char = SCCharacterizePeaks$new(lipid_sample, out_file = here::here("lipid_sample.zip"))

sc_char$load_file()
sc_char$prepare_raw_data()
```

#### Check Frequency Model

And now we can check that our frequency model should work for the rest of the steps.

```{r check_frequency}
sc_char$check_frequency_model()
```

If the frequency model was wrong, then we should instantiate a new one with a different model.

```{r example_char2}
freq_model2 = c("a.freq" = 0, "x.freq" = -1, "y.freq" = -1/3)

sc_char2 = SCCharacterizePeaks$new(lipid_sample,
                                   frequency_fit_description = freq_model2)

sc_char2$load_file()
sc_char2$prepare_raw_data()

sc_char2$check_frequency_model()
```

Yeah, this model is definitely incorrect

#### Check Scans Kept

We can also check which peaks were excluded by examining the `scan_info` data.frame.

```{r check_scans}
scan_info = sc_char$scan_info()
head(scan_info)

scan_info %>%
  ggplot(aes(x = rtime, xend = rtime, y = 0, yend = tic, color = keep)) +
  geom_segment()
```

We can see that several scans were excluded by being too high in their retention time (rtime), and then two were outliers based on the statistics of the frequency square root coefficient.

If this isn't doing what we want, we can define a different scan-filtering function that does more along the lines of what we want. See the documentation about `get_scan_info()` using `?get_scan_info` to see what is generally available.

#### Run Peak Characterization

Once we are happy with how we have set up the `SCCharacterizePeaks` object, we can run the peak characterization.

```{r run_characterization, eval = FALSE}
# not run
sc_char$find_peaks()
```

#### Run Everything

If you've already examined the model, and have your function for filtering scans all set up, then you can run everything on your sample rather quickly using this simple command:

```{r run_everything, eval = FALSE}
# not run
sc_char = SCCharacterizePeaks$new("file.mzML", out_file = "file.zip")
sc_char$run_all()
```

### SCRaw

`SCRaw` is responsible for loading the raw data, holding per-scan metadata, extracting the scan-level m/z and intensity, converting them to frequency, and filtering any scans out.
Although the user should not have to interact with it much, and **normally** it is controlled by `SCCharacterizePeaks`, we cover it first here b/c it is useful to instantiate it at least once per experiment and verify the scan filtering conditions and the model for converting m/z to frequency.

The default frequency ~ mz model is defined as:

$$frequency = a + \frac{x}{mz} + \frac{y}{\sqrt{mz}} + \frac{z}{\sqrt[3]{mz}}$$

We encode this using the vector: 

```{r freq_vector}
freq_vector = c("a.freq" = 0, "x.freq" = -1, "y.freq" = -1/2, "z.freq" = -1/3)
freq_vector
```

We can see how this corresponds to the equation above if we transform the equation to be like this:

$$frequency = a + x \times mz^{-1} + y \times mz^{\frac{-1}{2}} + z \times mz^{\frac{-1}{3}}$$

Here we use the **-** to define that it should be a fraction, and the **1/2** exponent corresponds to the square root.

This is the **default model** that `SCRaw` is instantiated with.
But every time you work with a new dataset, you should verify the model being used is appropriate.

```{r check_model}
lipid_1 = SCRaw$new(system.file("extdata", "lipid_example.mzML", package = "ScanCentricPeakCharacterization"))
lipid_1$extract_raw_data()
lipid_1$predict_frequency()

lipid_1$check_frequency_model()
```

So this model looks right.

What if we mis-define the model?

$$frequency = a + x \times mz^{-1} + y \times mz^{\frac{-1}{3}}$$

```{r alt_model}
alt_model = c("a.freq" = 0, "x.freq" = -1, "y.freq" = -1/3)

lipid_2 = SCRaw$new(system.file("extdata", "lipid_example.mzML", package = "ScanCentricPeakCharacterization"),
                    frequency_fit_description = alt_model)
lipid_2$extract_raw_data()
lipid_2$predict_frequency()

lipid_2$check_frequency_model()
```

Here we can see in the residuals (mean_predicted) that we've obviously mis-defined the model.
It's not obvious from the means vs predictions that we've mis-defined it!
