---
title: "Full Peak Characterization"
author: "Robert M Flight"
output: rmarkdown::html_vignette
package: ScanCentricPeakCharacterization
bibliography: scpc.bib
vignette: >
  %\VignetteIndexEntry{Full Peak Characterization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ScanCentricPeakCharacterization)
```

So you want to do some peak characterization, using `ScanCentricPeakCharacterization`, hereafter shortened to SCPC.
If you haven't already, you might want to read the publication describing the motivation behind it [@flightScanCentricFrequencyBasedMethod2022].

First things to know:

* Is your data in one of the open formats that `mzR` and `MSnbase` support (mzML, mzXML, mzData)?
* Do you have multiple, non-chromatographic scans or acquisitions of data where the same M/Z range is covered?
* Is your data acquired in profile mode?
* Is your data from an ICR or orbitrap type of instrument (Fusion, Fusion Lumos, SolariX, etc)?

If the answers to the above are all true, then you should be OK going forward.
If any of them are not true, or your aren't sure, then feel free to contact the package author, Robert Flight <rflight79 at gmail.com>, or [file an issue](https://github.com/MoseleyBioinformaticsLab/ScanCentricPeakCharacterization/) on the GitHub repo of this package.

## Logging

## References
