# How to Use R6 Classes

Ideally I want to use R6 classes for the mzml data and peaks, so lets work
out how best to actually do that.

## Current Plan

### AnalyzeMS

Class that controls how an analysis is run. Has methods for loading data,
and peak finding (PeakFinder). The PeakFinder should be set on the generator.

### PeakFinder

Takes as input two pieces from the RawMS data, the raw data and the ranges to 
use.

### ZipMS

This is the overall data container for a zipped file of mass-spec data.

Initializer should have arguments for which peices are actually loaded, with 
the default being everything.

**Data slots**:

* metadata: a MetaData data member
* raw_ms: a RawMS data member
* peaks: a Peaks data member

**Methods**:

* initialize
* read
* write

### RawMS

Data container for the raw mass-spec data, should take two filehandle to read in
the data, one for the data, and one for metadata.

**Data slots**:

* metadata:
* data:
* use_scans:

**Methods**:

* initialize
* set_scans


```{r load_up}
library(SIRM.FTMS.peakPickingMethods)
```

```{r set_files}
mzml_file <- "~/Projects/work/SIRM.FTMS.automatedAssignmentProcessExampleData/UK001N1exoposb.mzML"
zip_file <- "~/Projects/work/SIRM.FTMS.automatedAssignmentProcessExampleData/UK001N1exoposb.zip"
```

```{r test_mzml_class}
test1 <- ZipMS$new(zip_file)

test2 <- ZipMS$new(mzml_file)

test3 <- ZipMS$new(zip_file, out_file = "crap.txt")
```

```{r find_peaks}
find_peaks_default <- function(data, scan_range){
  avg_data <- as.data.frame(xcms::getSpec(data, scanrange = scan_range))
  avg_data$ObservedMZ <- avg_data$mz
  avg_data$Intensity <- avg_data$intensity
  avg_data$mz <- NULL
  avg_data$intensity <- NULL
  avg_peaks <- find_peaks_diff(avg_data)
  
  function_call <- "find_peaks_diff"
  function_pkg <- find(function_call)
  pkg_desc <- utils::packageDescription(substring(function_pkg, 9))
  if (!is.null(pkg_desc$RemoteSha)) {
    pkg_sha <- pkg_desc$RemoteSha
  } else {
    pkg_sha <- ""
  }
  pkg_version <- packageVersion(substring(function_pkg, 9))
  picking_description <- list(package = function_pkg,
                             version = pkg_desc$Version,
                             sha = pkg_sha,
                             function_called = function_call,
                             parameters = list(scan_range = force(scan_range))
                             )
  PeakPickingAnalysis$new(in_peaks = avg_peaks,
                          in_parameters = list(scan_range = force(scan_range),
                                                        picking_description = picking_description))
}
```


```{r try_it}
anal_ms <- AnalyzeMS$new(zip_file, file.path(getwd(), "UK001N1exoposb_out.zip"))
anal_ms$load_file()

anal_ms$set_peak_finder(find_peaks)
anal_ms$find_peaks()

anal_ms$zip_ms$out_file

anal_ms$zip_ms$save()
anal_ms$zip_ms$cleanup()
```

